---
title: "sentiment analysis ML"
author: "Alex Lin"
date: "2023-02-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

library(tidyverse)
library(tidytext)
library(dplyr)
library(tm)
library(textstem)
library(splitTools)
library(caret)

library(naivebayes)
library(e1071)
library(randomForest)
library(rpart)

```


## Data Processing: 

```{r}

congress_sample <- congress_sample[,c("date", "title", "text", "senti_label")]
pres_sample <- pres_sample[,c("date", "title", "text", "senti_label")]

us_sample <- rbind(pres_sample, congress_sample) %>% 
  drop_na(senti_label)

```


## Text Processing: 

A function that removes stopwords, lemmatizes the data, creates a document-term matrix, and performs tf-idf
```{r}

vectorizer_tfidf <- function (df) {
  
  corpus <- Corpus(VectorSource(df$text)) %>%  
  tm_map(content_transformer(function(x) removeWords(x, stopwords("en")))) %>% # removes stopwords
  tm_map(lemmatize_strings) # lemmatizes words

  corpus_dtm <- DocumentTermMatrix(corpus) %>% # creates document-term matrix
    weightTfIdf() %>% # tf-idf
    as.matrix() 
  
  # df_tfidf <- cbind(as.data.frame(corpus_dtm), senti_label = df$senti_label) 
  
  zero_var_cols <- nearZeroVar(corpus_dtm, saveMetrics = TRUE)$nzv
  corpus_dtm <- corpus_dtm[,-zero_var_cols]
  
  return(corpus_dtm)

}

```

Applying the function to US sample data: 
```{r}
us_sample_tfidf <- vectorizer_tfidf(us_sample)
```

## Dimensionality Reduction

Very crude dimensionality reduction by dropping very infrequent words:
```{r}
us_sample_tfidf <- us_sample_tfidf[,colSums(us_sample_tfidf != 0)>1]
```

Dimensionality reduction with PCA:
```{r}
pca.us <- prcomp(us_sample_tfidf, scale = T)
us_sample_tfidf <- pca.us$x[,1:100]
```

## Getting the data ready for ML

A function that performs train-test split:

```{r}
ml_ready <- function (dtm, original_data) {
  
  #Train-Test Split
  group_index <- partition(original_data$senti_label, p = c(train = 0.8, test = 0.2))

  train_data <- dtm[group_index$train,]
  test_data <- dtm[group_index$test,]
  
  #Converting dependent var to numeric
  original_data$senti_label_num <- ifelse(original_data$senti_label=="P", 1, -1)

  train_label <- original_data$senti_label_num[group_index$train]
  test_label <- original_data$senti_label_num[group_index$test]
  
  ml_list <- list("train_data" = train_data, "test_data" = test_data, 
                  "train_label" = train_label, "test_label" = test_label,
                  "group_index" = group_index)
  return(ml_list)
}
```

Applying the function to US data:
```{r}

ml_us <- ml_ready(us_sample_tfidf, us_sample)

train_us <- ml_us$train_data
test_us <- ml_us$test_data
train_us_label <- ml_us$train_label
test_us_label <- ml_us$test_label
us_group <- ml_us$group_index

```


## Naive Bayes

```{r}

us.nb <- naive_bayes(x=train_us, y = as.factor(train_us_label))
us.predict <- predict(us.nb, test_us)

us_sample_test <- us_sample[us_group$test,] 
us_sample_test$nb.label <- us.predict

```

Creating a confusion matrix for NB:
```{r}
table(us_sample_test$nb.label, test_us_label)
```

## SVM

```{r}

us.svm <- svm(x=train_us, y=as.factor(train_us_label), kernel = "linear")
us.predict <- predict(us.svm, test_us)

# us_sample_test <- us_sample[us_group$test,] 
us_sample_test$svm.label <- ifelse(us.predict == "1", 1, -1)

```

Creating a confusion matrix for SVM:
```{r}
table(us_sample_test$svm.label, test_us_label)
```
## Random Forest

```{r}

us.randomforest <- randomForest(x=train_us, y=as.factor(train_us_label),
                                mtry=25, tree=25)
us.predict <- predict(us.randomforest, test_us)

us_sample_test$randforest.label <- us.predict
```

Creating a confusion matrix for Random Forest:
```{r}
table(us_sample_test$randforest.label, test_us_label)
```

```{r}

us.dt <- rpart(formula = as.factor(train_us_label) ~., data = as.data.frame(train_us))
us.predict <- predict(us.dt, as.data.frame(test_us))

us_sample_test$dt.label <- ifelse(us.predict[,1] < us.predict[,2], 1,-1)
```

Creating a confusion matrix for Random Forest:
```{r}
table(us_sample_test$dt.label, test_us_label)
```


