---
title: "Time Series Analysis"
author: "Alex Lin"
date: "2023-03-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Time Series Analysis

Importing packges
```{r}

library(tseries)
library(lmtest)

```


## Processing US data: 
```{r}
# us_data <- rbind(pres.senti, congress.senti[,c(2,1,3,5,4,6,7,8,9,10)]) %>% drop_na(date)
us_data <- pres.senti %>% drop_na(date)
```

Creating a function that groups data by MONTH: 
```{r}
group_month <- function (df) {
  
  df$month <- format(df$date, "%Y-%m")
  
  df_month <- df %>%
    group_by(month) %>%
    summarise(mean_sentiment = mean(hybrid), doc_count = n())
  
  df_month$month <- as.Date(paste(df_month$month, "-01", sep = ""))
  
  return (df_month)
}

```

Creating a function that groups data by WEEK: 
```{r}
group_week <- function (df) {
    df$week <- difftime(df$date, as.Date("2000-01-01"), units = "weeks") %>%
    as.numeric() %>%
    floor()
  
    df_week <- df %>%
      group_by(week) %>%
      summarise(mean_sentiment = mean(hybrid), doc_count = n())
  
    df_week$week <- as.Date("2000-01-01") + df_week$week * 7
    
    return(df_week)
}
```


Applying to US data: 
```{r}
us_month <- group_month(us_data)
us_week <- group_week(us_data)
```

## Processing Chinese data: 
```{r}
mofa_month <- mofa.senti[!is.na(mofa.senti$date),] %>% group_month()
mofa_week <- mofa.senti[!is.na(mofa.senti$date),] %>% group_week()
```

## Plotting the Time Series
```{r}

ggplot() + geom_line(data = mofa_week, aes(x = week, y = mean_sentiment), color = "red", alpha = 0.1) +
  geom_ma(data = mofa_week, aes(x = week, y = mean_sentiment), color = "red", ma_fun = SMA) + 
  geom_line(data = us_week, aes(x = week, y = mean_sentiment), color = "darkblue", alpha = 0.1) +
  geom_ma(data = us_week, aes(x = week, y = mean_sentiment), color = "blue", ma_fun = SMA) + 
  xlab("date") + ylab("mean sentiment") + 
  theme_minimal()

```

## Granger Causality Assumptions

Checking whether US data is stationary with Dickey-Fuller
```{r}
adf.test(us_month$mean_sentiment)
```
Checking whether China data is stationary with Dickey-Fuller
```{r}
adf.test(mofa_month$mean_sentiment)
```

Because the MoFA data is not stationary, we will use first-order differencing to stabilize the mean: 
```{r}
mofa_month_diff <- diff(mofa_month$mean_sentiment)
us_month_diff <- diff(us_month$mean_sentiment)
```

Creating new master dataset with stationary data from both US and China: 
```{r}

mofa_month_2 <- mofa_month[-1,]
mofa_month_2$senti_diff <- mofa_month_diff

us_month_2 <- us_month[-1,]
us_month_2$senti_diff <- us_month_diff

time_series <- left_join(mofa_month_2, us_month_2, by = "month")
colnames(time_series) <- c("date","senti_china","ndoc_china","sentidiff_china",
                            "senti_us","ndoc_us","sentidiff_us")

```

Plotting first-order differences: 
```{r}

ggplot(time_series) + geom_line(aes(x = date, y = sentidiff_china), color = "red") +
  geom_line(aes(x = date, y = sentidiff_us), color = "darkblue") +
  xlim(as.Date(c('2019-01-01', '2022-01-01')))

```

## Performing Granger Causality!!

```{r}

gct <- grangertest(sentidiff_us ~ sentidiff_china, order = 1, 
                   data = time_series)
gct

```


